Title of Paper

### 2017-02-14 paper.tex
more edits

Several recent works, e.g. \cite{CurtisGPU:2017,Sewerin20151375} have investigated SIMT-based chemical kinetic integration on GPUs, however SIMD-based chemical kinetics have been less studied in comparison.
\textcite{stone2016} implemented a linearly-implicit fourth-order stiff Rosenbrock solver in the OpenCL for various platforms including the CPU, GPU and the MIC; the Rosenbrock solver performance was compared to that of a non-stiff, fourth-order explicit Runge–Kutta–Fehlberg integrator.
Two parallelization models were investigated in OpenCL, the so called \textit{shallow}-SIMD vectorization, where each SIMD-lane solved a single chemical kinetic initial value problem (IVP), and a SIMT-vectorization where each thread solved a single IVP.
SIMT-vectorized chemical source term evaluation was \SI{13}{\percent} and \SI{35}{\percent} slower than the equivalent OpenMP code on the CPU and MIC respectively.
The SIMD-vectorized chemical source terms were up to \SI{3}{$\times$} faster on both platforms than the OpenMP baseline; GPU chemical source term evaluation was of similar performance to the SIMD-vectorized versions.
SIMD-vectorization improved the integrator performance over the OpenMP baseline by \SIrange{2.5}{2.8}{$\times$} on the CPU and \SIrange{4.7}{4.9}{$\times$} on the MIC.
\textcite{kroshko2013efficient} implemented a shallow-SIMD -vectorized third order Rosenbrock solver (RODAS) integrator for atmospheric chemistry on a Cell Broadband Engine---a specially designed vector processor---finding a speedup of \SI{1.89}{$\times$} over a serial version of the same code, a \SI{94}{\percent} parallel efficiency.

This work will study the performance of chemical source term evaluation of automatically generated shallow-SIMD-vectorized codes for a wide range chemical kinetic models on the CPU.
The performance will be compared to a baseline SIMT-vectorized codes to determine the effective SIMD speedup.
Finally future extensions to this work will be detailed.
\todo[inline]{and OpenMp, time permitting}

@@

\subsection{Code Generation}
Code generation is handled by the python package \texttt{loo.py}~\cite{kloeckner_loopy_2014}, which provides an interface to OpenCL, while allowing for unit testing and simple changes of program structure, e.g. data ordering, vectorization, threading patterns etc.
Additionally, \texttt{loo.py} allows for code generation for multiple languages, including (non-vectorized) C, CUDA, and Intel's ISPC compiler, which targets SIMD vectorization on the CPU and MIC.
This is a direction that may be explored in the future to compare the speedups of the various compilers.

\subsection{Data Ordering and Vectorization Patterns}
For a chemical kinetic model with species $k$ for the $j$-th thermo-chemical state to be evaluated, $[C_k]_j$ there are two choices for data-storage in memory.
In the ``C'' (row-major) format, species concentrations for $j$-th thermo-chemical state are stored sequentially in memory, i.e.: $[C_0]_0\text{,}\allowbreak C[1]_0\ldots\allowbreak [C_{N_S}]_0\text{,}\allowbreak [C_0]_1\allowbreak\ldots\allowbreak [C_{N_S}]_N$, where $N$ is the total number of thermo-chemical conditions.
Alternatively, in the ``F'' (Fortan column-major) format species concentrations a single species for all thermo-chemical states are sequential, i.e.: $[C_0]_0\text{,}\allowbreak[C_0]\ldots\allowbreak[C_0]_N\text{,}\allowbreak[C_1]_0\ldots\allowbreak[C_{N_S}]_N$
In Sec.~\ref{S:results} the effect of this ordering will be shown to have a large performance impact on the program runtime.

>>>>

Several recent works, e.g. \cite{CurtisGPU:2017,Sewerin20151375} have investigated SIMT-based chemical kinetic integration on GPUs, however SIMD-based chemical kinetics have been less studied in comparison.
\textcite{stone2016} implemented a linearly-implicit fourth-order stiff Rosenbrock solver in the OpenCL for various platforms including the CPU, GPU and the MIC; the Rosenbrock solver performance was compared to that of a non-stiff, fourth-order explicit Runge–Kutta–Fehlberg integrator.
Two parallelization models were investigated in OpenCL, the so called \textit{shallow}-SIMD vectorization, where each SIMD-lane solved a single chemical kinetic initial value problem (IVP), and a SIMT-vectorization where each thread solved a single IVP.
Shallow-SIMD vectorization improved the integrator performance over the OpenMP baseline by \SIrange{2.5}{2.8}{$\times$} on the CPU and \SIrange{4.7}{4.9}{$\times$} on the MIC, while the GPU performance while the GPU was \SIrange{1.4}{1.6}{$\times$} slower than the OpenMP baseline due to thread-divergence concerns.
\textcite{kroshko2013efficient} implemented a shallow-SIMD -vectorized third order Rosenbrock solver (RODAS) integrator for atmospheric chemistry on a Cell Broadband Engine---a specially designed vector processor---finding a speedup of \SI{1.89}{$\times$} over a serial version of the same code, a \SI{94}{\percent} parallel efficiency.

This work will study the performance of chemical source term evaluation of automatically generated shallow-SIMD-vectorized codes for a wide range chemical kinetic models on the CPU.
The performance will be compared to a baseline SIMT-vectorized codes to determine the effective SIMD speedup.
Finally future extensions to this work will be detailed.

@@

\subsection{Code Generation}
Code generation is handled by the python package \texttt{loo.py}~\cite{kloeckner_loopy_2014}, which provides an interface to OpenCL allowing for unit testing and changes of program structure, e.g. data ordering, vectorization, threading patterns etc.
Additionally, \texttt{loo.py} allows for code generation for multiple languages, including (non-vectorized) C, CUDA, and Intel's ISPC compiler, which targets SIMD vectorization on the CPU and MIC.
This is a direction that may be explored in the future to compare the speedups of the various compilers.

\subsection{Data Ordering and Vectorization Patterns}
For a chemical kinetic model with species $k$ for the $j$-th thermo-chemical state to be evaluated, $[C_k]_j$ there are two choices for data-storage in memory.
In the ``C'' (row-major) format, species concentrations for $j$-th thermo-chemical state are stored sequentially in memory, i.e.: $[C_0]_0\text{,}\allowbreak C[1]_0\ldots\allowbreak [C_{N_S}]_0\text{,}\allowbreak [C_0]_1\allowbreak\ldots\allowbreak [C_{N_S}]_N$, where $N$ is the total number of thermo-chemical conditions.
Alternatively, in the ``F'' (Fortan column-major) format species concentrations a single species for all thermo-chemical states are sequential, i.e.: $[C_0]_0\text{,}\allowbreak[C_0]_1\ldots\allowbreak[C_0]_N\text{,}\allowbreak[C_1]_0\ldots\allowbreak[C_{N_S}]_N$
In Sec.~\ref{S:results} the effect of this ordering will be shown to have a large performance impact on the program runtime.
